---
title: "R Practice 13: Data Modeler"
author: "Ronald W. Buie"
format: pdf
---
                                                                                                                                                                          
*Render this document in r studio to see a pdf (or other) format*

Data modeler is a function in the [rads](https://github.com/PHSKC-APDE/rads) package that is designed to create uncorrelated versions of any two dimensional data frame or data table object. You can read more about it in the [wiki](https://github.com/PHSKC-APDE/rads/wiki/utilities#data_modeler) or in the man page by calling ?rads::data_modeler within R.


```{r setup}
library(rads)
library(apde.data)
library(data.table)
```

`data_modeler()` receives a data.table object and returns a data.table of similar structure and content, but with uncorrelated data across columns. For example, a column of dates will become a column of dates of a similar range, and distribution, but randomly distributed across observations. Common use cases include:

-   demonstrating a statistic or visualization without exposing the audience to "real" data or results
-   sharing code that generates a data set for minimally viable examples used in requesting help from, or teaching, others
-   creating data sets programmatically for (unit) testing or training

There are two kinds of output that `data_modeler()` returns, either the data table itself, or a list object of strings that can be evaluated to create the data table. In addition, `data_modeler()` can output-to-terminal a human readable version of the generation code that the user can copy and paste into their own scripts for sharing publicly.

`data_modeler()` makes some guesses as to if a column's content should be randomized (such as person names or blood pressures) or kept but randomly distributed (such as levels of response to a question or city names). This is intended to generate random data or to retain original categories where appropriate. It bases these guesses on the data type (integer or string), and complexity (e.g. 75% being unique data), but not the meaning of the data (i.e. it doesn't know that a number is a "zip code" and doesn't know that a string is a "human name"). Variables that are identified as non-categorical will, in most cases, be replaced with a uniform distribution of noisy, but similarly shaped, data (e.g. floating point numbers of the same range and precision, but in a random uniform distribution, or a "lore ipsum" string of similar lengths). Variables identified as categorical will, in most cases, be replaced with identical categories in a similar distribution to the original, but randomly assigned across observations (a data set of 95% males and 5% females would still generate a data set of roughly 95% males and 5% females).

Two parameters are available to adjust the assumptions data_modeler uses to identify categorical values. These are "categorical_threshold" and "proportional_threshold".

-   "categorical_threshold": defines how many unique categories can exist in the variable before data_modeler assumes the variable isn't categorical. Above this number, and the value will not be considered categorical. This currently defaults to 61.
-   "proportional_threshold": defines the proportion of values that can be unique before data_modeler() assumes the variable isn't categorical. This currently defaults to 0.75.

### data privacy caveats

#### uncorrelated may not always equal non-informative

`data_modeler()` creates an uncorrelated column of data to substitute the original by randomly reassigning values across that column. However, this does not ensure that meaningful information is not passed along. Note that, in most cases, categorical observations are distributed in proportion to the source data. So, if the provided data set is of categorical outcomes in a single population, the uncorrelated result would still show a similar rate of that outcome in that population because all observations would be for the same population group. The analyst should consider if this is acceptable or not and take measures to correct if necessary by either changing the input data or modifying the synthetic data. A simple approach in the above example would be to add one or more additional population groups so that the outcomes would be distributed removing the ability to estimate the rate for the smaller population (in an uncorrelated result, the rate of the categorical outcome would tend to be similar in all populations.)

### walkthrough

For the examples below, we create a "real" data set and show how to use `data_modeler()` and how the synthetic and "real" data differ.

Our demo "real" data has one indicator variable that represents an outcome that is never in the White population, sometimes in the Asian population, and always in other Races. In an uncorrelated (and so non-informative) synthetic data set, we would expect this indicator to no longer be biased by race.

```{r create demonstration "real" data}
DT.real.data <- data.table(
   id = 1:1000,
   chi_geo_kc = sample(c('King County',NA), 1000, replace = T),
   chi_race_7 = factor(sample(c("Asian", "AIAN", "Black", "Hispanic", "NHPI", "White", "Other", "Multiple", NA), 1000, replace = T, prob = c(.19,.01,.07,.11,.01,.35,.07,.14,.02)), levels = c("Asian", "AIAN", "Black", "Hispanic", "NHPI", "White", "Other", "Multiple", NA)),
   chi_sex = as.factor(sample(c("Male","Female"), 1000, replace = T)),
   chi_geo_region = factor(sample(c("South", "North", "Seattle", "East"), 1000, replace = T), levels = c("South","North","Seattle","East")))

# Attach a biased indicator.
#indicator1, all White population are "never", all Asians population, "sometimes", and everyone else, "always"
   DT.real.data[, indicator1 := ifelse(chi_race_7 == "White", "never",
                         ifelse(chi_race_7 == "Asian", "sometimes",
                                "always"))]
   
DT.real.data <- apde.data::hys()
```

```{r create a data.table of synthetic data}
#| echo: true


##Observe that modeled data have the same structure as the original
DT.synthetic.data <- rads::data_modeler(ph.data = DT.real.data, #our "real" data
                                number_of_observations =  100, #number of observations to create in the synthetic data
                                comments = TRUE, #whether or not to include comments in code (only matters for return_code == TRUE or print_code == TRUE)
                                return_code = FALSE, #return a list object of strings for creating the synthetic data instead of the data itself. leave FALSE to receive synthetic data
                              print_code = FALSE) #print the generation code to terminal. useful for both review and if you wish to copy into other contexts, such as a shareable script or in a chat with a LLM/AI

DT.synthetic.data <- rads::data_modeler(ph.data = DT.real.data, #our "real" data
                                number_of_observations =  100, #number of observations to create in the synthetic data
                                comments = TRUE, #whether or not to include comments in code (only matters for return_code == TRUE or print_code == TRUE)
                                return_code = FALSE, #return a list object of strings for creating the synthetic data instead of the data itself. leave FALSE to receive synthetic data
                              print_code = FALSE) 
```

You can observe that the data structures are the same for both data frames:

```{r}
#| echo: fenced
str(DT.real.data)
str(DT.synthetic.data)
```

But we can see that the original bias of White's in the never group, Asians in the sometimes group, and all others in the always group now no longer holds.

```{r}
#| echo: true
prop.table(table(DT.real.data[,.(indicator1,chi_race_7)]))
prop.table(table(DT.synthetic.data[,.(indicator1,chi_race_7)]))

```

Notice that you may see extreme, apparently biased, results in the outcomes. This is still randomly distributed and you can observe these change if you recreate the synthetic.

You can review, share, and execute the code used to create the synthetic data in two ways. Manually by printing it to terminal and copying it into your R script, or by receiving and processing the text object. The first is more useful for review and sharing with others, the later is more useful for programmatic handling of the instruction set.

```{r}
#| echo: true

DT.synthetic.data <- data_modeler(ph.data = DT.real.data,
                                number_of_observations =  100, 
                                comments = TRUE, 
                                return_code = FALSE,
                              print_code = TRUE) # change this to true to copy code from your terminal output
```

The code below is what is used to generate the synthetic data. Using this approach you can see the data, sample approach and rates for each variable. Because we ran `data_modeler()` with `comments = TRUE`, we also see a human-readable string at the end of each line describing how each variable was classified by data_modeler. This allows you to review the logic, make changes, and share the code with others.

```{r code from terminal}
DT <- data.table( 
`id` = as.integer(1:100), # as an identifying (all unique) integer 
`chi_geo_kc` = sample(c(NA, "King County"), 100, replace = TRUE, prob = c(0.494, 0.506)), # as categorical character (non factor) 
`chi_race_7` = factor(sample(c("Black", "White", "AIAN", NA, "Asian", "Multiple", "Other", "NHPI", "Hispanic"), 100, replace = TRUE, prob = c(0.193, 0.014, 0.067, 0.11, 0.012, 0.36, 0.064, 0.157, 0.023)), levels = c("Asian", "AIAN", "Black", "Hispanic", "NHPI", "White", "Other", "Multiple"), ordered = FALSE), # as factor 
`chi_sex` = factor(sample(c("Male", "Female"), 100, replace = TRUE, prob = c(0.518, 0.482)), levels = c("Female", "Male"), ordered = FALSE), # as factor 
`chi_geo_region` = factor(sample(c("Seattle", "South", "North", "East"), 100, replace = TRUE, prob = c(0.241, 0.229, 0.258, 0.272)), levels = c("South", "North", "Seattle", "East"), ordered = FALSE), # as factor 
`indicator1` = sample(c("always", "never", NA, "sometimes"), 100, replace = TRUE, prob = c(0.424, 0.36, 0.193, 0.023))) # as categorical character (non factor)
```

Note, the above approach requires the user to copy and paste output from their terminal. If you wish to check, modify, and extract the code without human intervention, you need to receive the text as a list object using `return_code = TRUE`:

```{r}

DT.synthetic.data.instructions <- data_modeler(ph.data = DT.real.data,
                                number_of_observations =  100, 
                                comments = FALSE, 
                                return_code = TRUE, #setting this to "TRUE" will cause data_modeler() to return a list object of the text instead of a DT of synthetic data
                                print_code = FALSE)

head(DT.synthetic.data.instructions)
```

After processing the text as desired, you can execute it, creating the desired synthetic data object, as follows:

```{r}
## Manually convert received code into data
codeText <- paste(unlist(DT.synthetic.data.instructions), collapse =" \n" )
DT.synthetic.data.manual <- eval( parse(text = paste0(codeText)))
```

Now you can visually confirm that the class structure of each of the tables is the same:

```{r}
str(DT.real.data)
str(DT.synthetic.data)
str(DT.synthetic.data.manual)

sapply(DT.real.data, class)
sapply(DT.synthetic.data, class)
sapply(DT.synthetic.data.manual, class)

```

Finally, each time data_modeler is executed, there is a randomness to how the data are created. If you wish to create identical data sets, be sure to set seed before executing:

```{r}
DT.synthetic.data.random <- data_modeler(ph.data = DT.real.data,
                                number_of_observations =  100, 
                                comments = TRUE, 
                                return_code = FALSE,
                                print_code = FALSE) 

set.seed(1000)
DT.synthetic.data.s1000a <- data_modeler(ph.data = DT.real.data,
                                number_of_observations =  100, 
                                comments = TRUE, 
                                return_code = FALSE,
                                print_code = FALSE) 

set.seed(1000)
DT.synthetic.data.s1000b <- data_modeler(ph.data = DT.real.data,
                                number_of_observations =  100, 
                                comments = TRUE, 
                                return_code = FALSE,
                                print_code = FALSE) 

```

Comparing a table with a random seed to one with a fixed seed, we expect the resulting tables to differ:

```{r}
identical(DT.synthetic.data.random, DT.synthetic.data.s1000a)
```

When using the same seed for distinct rounds, we would expect the resulting tables to be identical:

```{r}
identical(DT.synthetic.data.s1000a, DT.synthetic.data.s1000b)

data_modeler()
```

Performance of the data typing heuristics can be adjusted by two parameters, "categorical_threshold" and "proportion_threshold". If you find something is being classified as a categorical when it should be free-text or continuous, or are finding the opposite, adjusting these thresholds might be helpful.

categorical_threshold : Threshold for the number of unique values in a variable at or below which data_modeler() will treat the value as cateogircal, default 61.

proportional_threshold	: Threshold for the proportion of observations that must be unique, below which data modeler() will treat the value as categorical, default 0.75

In general, both definitions must be met for a variable to be classified as categorical or not.

*mot all data types use these, and they are still a work-in-progress*
