---
title: "Spot check data"
author: "Alastair Matheson (Assessment, Policy Development, and Evaluation Unit)"
date: "9/11/2019"
output: 
  # md_document: default
  # keep_md: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

## Set up

1) Load the libraries used in this tutorial

```{r loadpacks}
# Importing data
library(data.table)

# Data manipulation
library(tidyverse)

```

2) Set up path to where the training data live
* This makes the assumption that your working directory is the Github repo R_training.

```{make_path}
data_path <- file.path(getwd(), "dummy_data")
```

3) Bring in some data

```{import_data}
training_data_xl_boys <- read_xlsx(file.path(data_path, "r_training_data.xlsx"),
                                   sheet = "boys_names", skip = 3)
```


## Data names and structure
### Data names
```{data_names}
# See the names of each field in the data
# Note that this imported data has messy names
names(training_data_xl_boys)

# Rename using dplyr's rename function
# Format is new_name = old_name
# Note use of `backticks` around names with invalid characters or spaces
training_data_xl_boys <- training_data_xl_boys %>%
  rename(rank = Rank...1,
         first_name = `First Name...2`,
         n = N...3,
         pct = `%...4`,
         cum_n = N...5,
         cum_pct = `%...6`)
```

### Data structure
```{data_structure}
# See the types of each field in the data
str(training_data_xl_boys)

# Look at the first few rows of data (default is 6 rows)
# Also shows the data type of each column
head(training_data_xl_boys)
# Show more rows
head(training_data_xl_boys, n = 10)

# Look at the last few rows of data
tail(training_data_xl_boys)
tail(training_data_xl_boys, 10)
```

### Look for columns with only missing data
* A few ways to approach this.
* The code below shows each field and the percent missing.
* It uses summarise_all from the dplyr package.

```{all_missing}
summarise_all(training_data_xl_boys, list(~sum(is.na(.)) / nrow(training_data_xl_boys)))
```

### Drop unwanted columns
* Use dplyr's select function.
* Can use it to specify desired columns or columns to drop

```{drop_cols}
# Drop a specific column with the - prefix
training_data_xl_boys <- training_data_xl_boys %>%
  select(-`...7`)

# Specify columns to include. Ranges can be shown with a :
training_data_xl_boys <- training_data_xl_boys %>%
  select(rank, first_name:cum_pct, N...12)

```
