---
title: "Medicaid claims 101"
author: "Eli Kern | PHSKC-APDE |"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome

This script serves as a companion to the Medicaid claims 101 training slide deck developed by PHSKC-APDE for PHSKC and DCHS data analysts. The [training slide deck](https://kc1.sharepoint.com/:p:/r/teams/DPH-KCCross-SectorData/Shared%20Documents/Training/PHSKC_Medicaid%20claims%20data%20101.pptx?d=w1eed7465cdcc49838dc0afd54d62a97c&csf=1&web=1&e=3ob1Yr) can be accessed by members of the [DPH-KCCross-SectorData](https://kc1.sharepoint.com/teams/DPH-KCCross-SectorData) SharePoint site.

This script is divided into sections corresponding to topics featured in the training slide deck.

## Setup the environment

First, some code to set options, load package libraries, and establish a connection to Medicaid claims data in the Health & Human Services Analyst Workspace (HHSAW):

```{r environment setup}
##### Set up global parameters and call in libraries #####
options(max.print = 350, tibble.print_max = 50, scipen = 999)
origin <- "1970-01-01" # Date origin
#devtools::install_github("PHSKC-APDE/claims_data") #install fresh package if it's been awhile
pacman::p_load(tidyverse, odbc, rads, openxlsx2, claims, rlang, glue, keyring)

#Connect to HHSAW using ODBC driver
db_hhsaw <- DBI::dbConnect(odbc::odbc(),
                           driver = "ODBC Driver 17 for SQL Server",
                           server = "tcp:kcitazrhpasqlprp16.azds.kingcounty.gov,1433",
                           database = "hhs_analytics_workspace",
                           uid = keyring::key_list("hhsaw")[["username"]],
                           pwd = keyring::key_get("hhsaw", keyring::key_list("hhsaw")[["username"]]),
                           Encrypt = "yes",
                           TrustServerCertificate = "yes",
                           Authentication = "ActiveDirectoryPassword")
```

## Counting distinct individuals in claims data

**Counting distinct people.**  
The following code counts the total number of people in Medicaid data across all time. PHSKC maintains Medicaid claims data for the most recent 10 years.

```{r all-time person count}

#Prep SQL query
sql_query_1 <- glue::glue_sql(
  "select count(distinct id_mcaid) as id_dcount
  from claims.final_mcaid_elig_demo;",
  .con = db_hhsaw)

#Send query to HHSAW, store result in R
result1 <- dbGetQuery(conn = db_hhsaw, statement = sql_query_1)
result1$id_dcount

```

**Counting people by mutually inclusive race/ethnicity categories.**  
Now let's count distinct individuals by race/ethnicity using a few different approaches. First, APDE's primary approach - mutually inclusive race/ethnicity categories:

```{r race/eth aic}

sql_query_1 <- glue::glue_sql(
  "select 'AI/AN' as race_eth, count(distinct id_mcaid) as id_dcount
  from claims.final_mcaid_elig_demo
  where race_aian = 1
  union select 'Asian' as race_eth, count(distinct id_mcaid) as id_dcount
  from claims.final_mcaid_elig_demo
  where race_asian = 1
  union select 'Black' as race_eth, count(distinct id_mcaid) as id_dcount
  from claims.final_mcaid_elig_demo
  where race_black = 1
  union select 'Latino' as race_eth, count(distinct id_mcaid) as id_dcount
  from claims.final_mcaid_elig_demo
  where race_latino = 1
  union select 'NH/PI' as race_eth, count(distinct id_mcaid) as id_dcount
  from claims.final_mcaid_elig_demo
  where race_nhpi = 1
  union select 'White' as race_eth, count(distinct id_mcaid) as id_dcount
  from claims.final_mcaid_elig_demo
  where race_white = 1
  union select 'Unknown' as race_eth, count(distinct id_mcaid) as id_dcount
  from claims.final_mcaid_elig_demo
  where race_eth_unk = 1;",
  .con = db_hhsaw)

result1 <- dbGetQuery(conn = db_hhsaw, statement = sql_query_1)
arrange(result1, race_eth)

```

**Counting people by mutually exclusive race/ethnicity categories.**  
Now let's count people using mutually exclusive race/ethnicity categories where people identifying with more than one race/ethnicity are included in the Multiple Race group.

```{r race_eth_me}

sql_query_1 <- glue::glue_sql(
  "select race_eth_me, count(distinct id_mcaid) as id_dcount
  from claims.final_mcaid_elig_demo
  group by race_eth_me;",
  .con = db_hhsaw)

result1 <- dbGetQuery(conn = db_hhsaw, statement = sql_query_1)
arrange(result1, race_eth_me)

```

**Calculating age.**  
Now let's calculate the minimum, maximum, and mean age of Medicaid beneficiaries. To calculate age, we have to choose a reference date - let's choose December 31, 2023.

```{r age}

reference_date <- "2023-12-31"

sql_query_1 <- glue::glue_sql(
  "with age_20231231 as (
  	select id_mcaid,
  	case
  		when (datediff(day, dob, {reference_date}) + 1) >= 0 then floor((datediff(day, dob, {reference_date}) + 1) / 365.25)
      when datediff(day, dob, {reference_date}) < 0 then NULL
    end as age
  	from claims.final_mcaid_elig_demo
  )
  select min(age) as age_min, max(age) as age_max, cast(round(avg(age),1) as numeric(4,1)) as age_mean
  from age_20231231;",
  .con = db_hhsaw)

result1 <- dbGetQuery(conn = db_hhsaw, statement = sql_query_1)
result1

```

**Selecting people with coverage during a measurement window.**  
Now let's start to work with time-varying concepts, beginning with coverage start and end dates. Note that people enter and exit the Medicaid data for three reasons: 1) people gain or lose Medicaid coverage (change in eligibility), 2) people move into or out of King County, or 3) people are born or die. With the exception of people entering due to birth, we cannot differentiate between any of the other reasons.

For this first example, let's find all the people with 1 or more day of Medicaid coverage in 2023. The `[claims].[final_mcaid_elig_timevar]` table contains from and to dates for a host of time-varying concepts - we need to create new from and to dates within our measurement window (2023-01-01 through 2023-12-31) in order to accurately count coverage days.

```{r overall_mcaid coverage cohort}

reference_from_date <- "2023-01-01"
reference_to_date <- "2023-12-31"

sql_query_1 <- glue::glue_sql(
  "with cov_2023 as (
    select id_mcaid, from_date, to_date, geo_kc, cov_time_day,
    
    case
    	--coverage period fully contains date range
    	when from_date <= {reference_from_date} and to_date >= {reference_to_date}
    	  then datediff(day, {reference_from_date}, {reference_to_date}) + 1
    	--coverage period begins before and ends within date range
    	when from_date <= {reference_from_date} and to_date < {reference_to_date}and to_date >= {reference_from_date}
    	  then datediff(day, {reference_from_date}, to_date) + 1
    	--coverage period begins within and ends after date range
    	when from_date > {reference_from_date}  and to_date >= {reference_to_date} and from_date <= {reference_to_date}
    	  then datediff(day, from_date, {reference_to_date}) + 1
    	--coverage period begins and ends within date range
    	when from_date > {reference_from_date} and to_date < {reference_to_date}
    	  then datediff(day, from_date, to_date) + 1
    	else 0
    end as custom_cov_time_day

    from claims.final_mcaid_elig_timevar
    where from_date <= {reference_to_date} and to_date >= {reference_from_date}
    and geo_kc = 1
  )
  select count(distinct id_mcaid)
  from cov_2023
  where custom_cov_time_day >= 1;",
  .con = db_hhsaw)

result1 <- dbGetQuery(conn = db_hhsaw, statement = sql_query_1)
result1

```

For the second example, let's find all  people with 50% or more of 2023 covered by Medicaid.

```{r 50% and above coverage cohort}

reference_from_date <- "2023-01-01"
reference_to_date <- "2023-12-31"
reference_period <- interval(ymd(reference_from_date), ymd(reference_to_date))
reference_days <- reference_period %/% days(1) + 1

sql_query_1 <- glue::glue_sql(
  "with cov_2023 as (
    select id_mcaid, from_date, to_date, geo_kc, cov_time_day,
    
    case
    	--coverage period fully contains date range
    	when from_date <= {reference_from_date} and to_date >= {reference_to_date}
    	  then datediff(day, {reference_from_date}, {reference_to_date}) + 1
    	--coverage period begins before and ends within date range
    	when from_date <= {reference_from_date} and to_date < {reference_to_date}and to_date >= {reference_from_date}
    	  then datediff(day, {reference_from_date}, to_date) + 1
    	--coverage period begins within and ends after date range
    	when from_date > {reference_from_date}  and to_date >= {reference_to_date} and from_date <= {reference_to_date}
    	  then datediff(day, from_date, {reference_to_date}) + 1
    	--coverage period begins and ends within date range
    	when from_date > {reference_from_date} and to_date < {reference_to_date}
    	  then datediff(day, from_date, to_date) + 1
    	else 0
    end as custom_cov_time_day

    from claims.final_mcaid_elig_timevar
    where from_date <= {reference_to_date} and to_date >= {reference_from_date}
    and geo_kc = 1
  )
  select count(distinct id_mcaid)
  from cov_2023
  where custom_cov_time_day*1.0/{reference_days}*100.0 >= 50.0;",
  .con = db_hhsaw)

result1 <- dbGetQuery(conn = db_hhsaw, statement = sql_query_1)
result1

```

**Assigning people to a single geography for a given measurement window.**  
People move and thus when conducting an analysis for a given measurement window (let's use 2023 again), we often want to assign each person only once to geographic concepts, such as ZIP code of residence, to avoid people being counted more than once in a descriptive analysis.

Let's use the following code to 1) assign a single King County ZIP code to each Medicaid beneficiary for 2023, and 2) count distinct people by ZIP code of residence.

```{r zip code selection}

reference_from_date <- "2023-01-01"
reference_to_date <- "2023-12-31"

sql_query_1 <- glue::glue_sql(
  "with cov_2023 as (
    select distinct id_mcaid, geo_zip,
    
    case
    	--coverage period fully contains date range
    	when from_date <= {reference_from_date} and to_date >= {reference_to_date}
    	  then datediff(day, {reference_from_date}, {reference_to_date}) + 1
    	--coverage period begins before and ends within date range
    	when from_date <= {reference_from_date} and to_date < {reference_to_date}and to_date >= {reference_from_date}
    	  then datediff(day, {reference_from_date}, to_date) + 1
    	--coverage period begins within and ends after date range
    	when from_date > {reference_from_date}  and to_date >= {reference_to_date} and from_date <= {reference_to_date}
    	  then datediff(day, from_date, {reference_to_date}) + 1
    	--coverage period begins and ends within date range
    	when from_date > {reference_from_date} and to_date < {reference_to_date}
    	  then datediff(day, from_date, to_date) + 1
    	else 0
    end as custom_cov_time_day

    from claims.final_mcaid_elig_timevar
    where from_date <= {reference_to_date} and to_date >= {reference_from_date}
    and geo_kc = 1
  ),
  
  zip_code_cov_time as (
    select id_mcaid, geo_zip, sum(custom_cov_time_day) as custom_cov_time_day
    from cov_2023
    group by id_mcaid, geo_zip
  ),
  
  zip_code_ranks as (
    select id_mcaid, geo_zip, custom_cov_time_day,
    rank() over(partition by id_mcaid
      order by case when geo_zip is null then 1 else 0 end, custom_cov_time_day desc, geo_zip)
	    as geo_zip_rank
    from zip_code_cov_time
  )
  
  select geo_zip, count(distinct id_mcaid) as id_dcount
  from zip_code_ranks
  where geo_zip_rank = 1
  group by geo_zip;",
  .con = db_hhsaw)

result1 <- dbGetQuery(conn = db_hhsaw, statement = sql_query_1)

#Apply small number suppression
result1_cat1 <- result1 %>%
  mutate(id_dcount = case_when(
    between(id_dcount, 1, 10) ~ NA_integer_,
    TRUE ~ id_dcount
  ))
arrange(result1_cat1, desc(id_dcount))
```

## Counting health care encounters

**Counting emergency department visits.**  
The following code counts the total number of *distinct* emergency department (ED) visits by year from 2018-2022, using two different definitions for ED visit. The standard definition used by APDE is the `ed_pophealth_id` field, as uses a broader definition of an ED visit than the `ed_perform_id` field.

```{r ed_visits}

sql_query_1 <- glue::glue_sql(
  "select a.service_year, a.ed_pophealth_dcount, b.ed_perform_dcount
  from (
    select year(last_service_date) as service_year, count(distinct ed_pophealth_id) as ed_pophealth_dcount
    from claims.final_mcaid_claim_header
    where year(last_service_date) between 2018 and 2022
      and ed_pophealth_id is not null
    group by year(last_service_date)
  ) as a
  left join (
    select year(last_service_date) as service_year, count(distinct ed_perform_id) as ed_perform_dcount
    from claims.final_mcaid_claim_header
    where year(last_service_date) between 2018 and 2022
      and ed_perform_id is not null
    group by year(last_service_date)
  ) as b
  on a.service_year = b.service_year;",
  .con = db_hhsaw)

result1 <- dbGetQuery(conn = db_hhsaw, statement = sql_query_1)
arrange(result1, service_year)

```

**Counting inpatient hospital stays.**  
The following code counts the total number of *distinct* acute inpatient stays by year from 2018-2022, which includes hospitalizations for both medical reasons and for labor and delivery.

```{r inpatient}

sql_query_1 <- glue::glue_sql(
  "select year(last_service_date) as service_year, count(distinct inpatient_id) as inpatient_dcount
  from claims.final_mcaid_claim_header
  where year(last_service_date) between 2018 and 2022
    and inpatient_id is not null
  group by year(last_service_date);",
  .con = db_hhsaw)

result1 <- dbGetQuery(conn = db_hhsaw, statement = sql_query_1)
arrange(result1, service_year)

```

**Counting primary care visits.**  
The following code counts the total number of *distinct* primary care visits by year from 2018-2022.

```{r primary_care_visit}

sql_query_1 <- glue::glue_sql(
  "select year(last_service_date) as service_year, count(distinct pc_visit_id) as pc_visit_dcount
  from claims.final_mcaid_claim_header
  where year(last_service_date) between 2018 and 2022
    and pc_visit_id is not null
  group by year(last_service_date);",
  .con = db_hhsaw)

result1 <- dbGetQuery(conn = db_hhsaw, statement = sql_query_1)
arrange(result1, service_year)

```